### 2019-04-02 papers/journal/encoding.tex
QBF encoding updated

%\subsection{Boolean satisfiability of the search problem}

We translate the search problem into a Boolean satisfiability
problem and use SMT solvers to find the satisfying VTSs.
%
We will first present the variables used to encode the
VTSs and the properties.
%
Afterwards, we will present the constraints corresponding to the
properties.

@@

%if all the pairs of the present molecule are present on the same edge than all of them inhibit each other resulting in making all the molecules inactive, we term this process ``pairing inhibition".
%
The constraints for the variants can be given to a SAT or a SMT solver to find
VTSs that belong to the variants.

\subsection{Encoding for $\fqbf$ search problem}
\label{enc:univ-cond}
%
Recall that the specification for universal condition~\ref{sec:property} requires quantifier alternation. 
%
Most SAT solvers can not efficiently handle the quantifier in the formulas. 
% 
Although, it is possible to encode the universal condition as a SAT or SMT problem.
%
There is always a danger of a blow-up in the formula size, making it harder for the SAT solver to handle. 
%Which may blow up the formula size for the solver to solve.
% 
As an experiment, we have implemented a direct SMT version of the universal problem extending the current $\fsmt$ search encodings.
%
The basic idea is to conjunct the formula for each combination value of the quantified variable. 
%
%
\input{satqbf-table}
%
\todo{Fix the table; the entries of constraint is incorrect}The table~\ref{tab:satqbf-graph}, shows the search results of this encoding for the universal condition with 4 nodes and 3-connected graphs. 
%
The SMT extended version was not able to scale it to the desired results and either timed-out (T/O) or ran out of memory (M/O).
%
This can be due to the fact that directly translating a quantified formula into a quantifier-free  by introducing a vector of variables for each boolean possibility will blow up the size of the formula and may not be solvable by a solver.
%

\textbf{$\fqbf$ search encoding of the universal condition:}
We will describe several constraints that encode VTSs in this section.
%
To avoid cumbersome notation, we will not explicitly write the ranges
of the indexing in the constraints.
%
$i$ and $j$ will range over nodes, i.e., from $1$ to $\nu$.
%
$m$ will range over molecules, i.e., from $1$ to $\mu$.
%
$q$ will range over edges between two nodes, i.e., from $1$ to $\pi$.
%

The following constraints encode the basic consistency of VTSs. 
%
Recall that the molecule set is divided into 3:1 Q and R pairs (or partition).
\begin{align*}
\texttt{EdgeC} =\;&\hspace{-1ex}\bigwedge\limits_{i,j,q} (\bigvee_m e_{i,j,q,m} )\limplies e_{i,j,q}
\land
\bigwedge\limits_{i,q} \neg e_{i,i,q}
\land
% \tag{\texttt{NodeC}}\label{eq:f0}\\
\bigwedge\limits_{\mathclap{i,j,q,m}} e_{i,j,q,m} \limplies (n_{i,m} \land n_{j,m} )
\\
\texttt{ActivityC} =\;&
\bigwedge\limits_{\mathclap{i,j,q,m}} b_{i,j,q,m} \limplies e_{i,j,q,m} \quad\land\quad
\bigwedge\limits_{i,m} a_{i,m} \limplies n_{i,m}
% \tag{V2}\label{eq:f1}\\
\\
\texttt{PairingC} =\;&
\exists qr. \Land\limits_{\mathrlap{m_1,m_2,m_3,m_4}\;\;}
(p_{\{m_1,m_2,m_3,m_4\}} \limplies qr_{m_1} + qr_{m_2} + qr_{m_3} + qr_{m_4} = 3) 
% \land\\ 
%   & \quad\quad\quad ( m_1 = m_2 = m_3 = m_4  \limplies \lnot p_{m_1,m_2,m_3,m_4} )
% \bigwedge\limits_{m} 
% (\Lor\limits_{m'} p_{mm'} \limplies \Land_{m'} \lnot p_{m'm} ) \land
% (\Lor\limits_{m'} p_{m'm} \limplies \Land_{m'} \lnot p_{mm'} )
\\
\texttt{Fusion1} =\;&
\bigwedge\limits_{i,j,q} e_{i,j,q} \limplies
\bigvee_{{m_1,m_2,m_3,m_4}} (\Land_{l=1}^4 ( b_{i,j,q,m_l} \lor a_{j,m_l} ) \land 
\Lor\limits_{l=1}^4 b_{i,j,q,m_l} \land\\
& \qquad \qquad \qquad \qquad \Lor\limits_{l=1}^4 a_{j,m_l}\land p_{\{m_1,m_2,m_3,m_4\}})
\\
\texttt{Fusion2} =\;&
\bigwedge\limits_{\mathclap{i,j,q,m_1,m_2,m_3, l \in \{1,..,3\}}} b_{i,j,q,m_1} \land ..\land b_{i,j,q,m_l} \limplies \\
& \hspace{2cm} \neg 
\bigvee_{\mathclap{j \neq j^{\prime}, m_{l+1}^{\prime},..,m_{4}^{\prime}}} ( a_{j^{\prime},m_{l+1}^{\prime}} \land .. \land a_{j^{\prime},m_4^{\prime}} \land p_{\{m_1,..,m_l,m^{\prime}_{l+1},..,m^{\prime}_4\}})
\\
\texttt{Consistancy} =\;& \texttt{EdgeC} \land
\texttt{ActivityC} \land \texttt{PairingC} \land
\texttt{Fusion1} \land \texttt{Fusion2} 
\end{align*}
\texttt{EdgeC} states that each edge has at least one molecule,
there are no self loops, and edge labels are consistent with node labels.   
\texttt{ActivityC} states that active molecule are present.
\texttt{PairingC} states that all molecules are divided into two types
using $qr_m$ bit, which encodes if $m$ belongs to one type or another,
and any fusing set of molecules must have three molecules involved
from one type and one molecule from the other.
\texttt{Fusion1}, and \texttt{Fusion2} states the well-fused condition.
\texttt{Consistancy} is the conjunction of all of the above.

\paragraph{Activity functions}
%
We also need to encode that the activity of the molecules are
controlled by activity functions.
%
The input VTS may include concrete activity functions for some molecules,
and for the others the functions may be unknown and to be synthesized. 
%
The concrete functions can be given to us in many different ways,
for example as a lookup table, or a concise Boolean formula.
%
In the following section, we will assume the appropriate encoding is
used for the concrete functions and represent them by \texttt{NodeFun}$_m$
and \texttt{EdgeFun}$_m$ for node and edge regulations respectively.
%
%We will use $\nodef_m$ and $\edgef_m$ to represent functions that
%are unknown in a VTS.
%
% \begin{align}
% \bigwedge\limits_{i,m} n_{i,m} \limplies a_{i,m} =  \nodef_m (n_{i,1},\dots,n_{i,\mu}) 
% \tag{\texttt{ActiveN}}\label{eq:anb}\\
%    \bigwedge\limits_{i,j,q,m} e_{i,j,q,m} \limplies b_{i,j,q,m} = \edgef_m(e_{i,j,q,1}, .., e_{i,j,q,\mu} )
%   \tag{\texttt{ActiveE}}\label{eq:aeb}
% \end{align}
%
%Later we will be synthesizing the unknown activity functions and 
%replace $\nodef_m$ and $\edgef_m$ with parameterized constraints that
%encode a space of candidate functions.
%\subsubsection{VTS properties}
%For the synthesis of incomplete systems,
%we need properties against which we synthesize the missing parts.
%
Next, we will discuss stability and $k$-connectedness.
%

\paragraph{Stability property}
%
We use Boolean variable $r_{i,j,m,p}$ to indicate if there is an
$m$-path from $i$ to $j$ of length less than or equal to $p$.
%
We use $m$-reachability to encode the stability condition in VTSs.
%
The following constraint recursively encodes that node $j$ is
$m$-reachable from node $i$ in less than $p$ steps.
%
Subsequently, we encode stability condition using the reachability variables.
\begin{align*}
\texttt{Paths}(r) &= \bigwedge\limits_{\mathclap{i,j,m,p}} r_{i,j,m,p} \limplies (\bigvee_{q} \, e_{i,j,q,m} \lor \bigvee_{i\neq i^{\prime}} ( \, \bigvee_{q} e_{i,i^{\prime},q,m}) \land r_{i^{\prime},j,m,p-1} )
\\
\texttt{Loop}(r) &= \bigwedge\limits_{i,j,m} (\bigvee_{q} e_{i,j,q,m}) \limplies r_{j,i,m,\nu}
\\
\texttt{Stability} &= \exists r. \; \texttt{Paths}(r) \land \texttt{Loop}(r)
\end{align*}

\paragraph{$k$-connected property}
%
$k$-connectedness expresses robustness against failure of few edges.
%
Let us use $d_{i,j,q}$ to indicate $q$th edge between $i$ and $j$ is failed
and $r'_{i,j}$ to indicate if there is a path from $i$ to $j$ in
the modified VTS.
%
% To check whether $k$-connected is a
% necessary condition, we remove (drop) $k-1$ edges from the graph and
% if it disconnects the graph, and we get an assignment.
% %
% We have a graph that is not $k$-connected.
%
In the following, $\texttt{Fail}(d,k)$ encodes that only
existing edges can be failed and exactly $k-1$ edges are failed.
%
$\texttt{FReach}(d,r')$ defines reachability in the modified VTS.
%
We use a new variable $r'_{i,j,p}$ to encode reachability from
$i$ to $j$ in at most $p$ steps.
%
$\texttt{Connected}(r')$ says that all nodes are reachable from any
other node.
\begin{align*}
\texttt{Fail}(d,k) = & 
\bigwedge\limits_{i,j,q} d_{i,j,q} \limplies e_{i,j,q}  \land 
\sum_{i,j,q} d_{i,j,q} = k-1\\
% \texttt{ReachAbove}(d,r') = &
% \bigwedge\limits_{i,j}  [\bigvee_{q} (e_{i,j,q} \land  \neg d_{i,j,q}) \lor  (\bigvee_{i' \neq i}  r^{\prime}_{i',j} \land  \bigvee_{q} (e_{i,i',q} \land \neg d_{i,i',q}) ] \limplies r^{\prime}_{i,j}\\
\texttt{FReach}(d,r') = &\hspace{-1ex}
\bigwedge\limits_{i,j,p}  \hspace{-1ex}r^{\prime}_{i,j,p} \hspace{-1ex}\limplies\hspace{-1ex} [\bigvee_{q} (e_{i,j,q} \land  \neg d_{i,j,q}) \hspace{-1pt}\lor \hspace{-2pt} (\bigvee_{\mathclap{i' \neq i}}  r^{\prime}_{i',j,p-1} \land  \bigvee_{q} (e_{i,i',q} \land \neg d_{i,i',q}) ]\\
\texttt{Connected}(r') = & \Land\limits_{i,j} (r^{\prime}_{i,j,\nu} \lor r^{\prime}_{j,i,\nu})
\end{align*}
%We will be synthesizing $k$-connected graphs.
%
We define $\texttt{Connected}(k)$ that says for all possible valid failures
the graph remains reachable. 
\begin{align*}
\texttt{Connected}(k) = & \forall d.\;
(\texttt{Fail}(d,k) \limplies \exists r'.\;\texttt{FReach}(d,r')
\land \texttt{Connected}(r'))
% \\
% \texttt{Disconnected}(k) = & \exists d.\;
%         (\texttt{Drop}(d,k) \land \exists r'.\;\texttt{ReachAbove}(d,r')
%                               \land \lnot \texttt{Connected}(r'))
\end{align*}
Since $d$ variables in $\texttt{Connected}(k)$ are universally
quantified, $\texttt{Connected}(k)$ introduces quantifier alternations.
%
%Therefore, synthesis against this property will require QBF reasoning.
%%
%We may make the formula quantifier free  by considering all possible failures
%separately and introducing a vector of reachability variables for each
%failure.
%%
%However, this will blow up the size of the formula and may not be
%solvable by a SAT solver.
%
%\begin{align*}
%  & \forall G \, \texttt{Connectivity(k)} \limplies  \exists
%                        f,p: \texttt{Stability} \, \land \, \texttt{Consistancy}  
%  \tag{U}\label{eq:univcond}
%\end{align*}

@@

In our synthesis method, we take a VTS $G =
(\nodes,\mols,\edges,\nlabel,\pairs,\edgef,\nodef)$ as input.
%
We allow activity functions not to be specified.
%
We will re-use the $\fqbf$ encodings for the input VTS as described in the Section~\ref{enc:univ-cond}.
%
We construct the following constraints to encode the available information
about $G$.
%
We encode both the present and the absent components in $G$.
%
Later, the constraints will help us encode the synthesis problems.
%
\begin{align*}
\texttt{PresentE} &= \land \{e_{i,j,q,m}|(i,M_1,j),...,(i,M_{q'},j) \in \edges \land q \leq q' \land m \in M_q \}\\
\texttt{PresentN} &= \land \{n_{i,m}| m \in \nlabel(i) \land i \in \nodes \}\\
\texttt{PresentP} &= \land \{p_{\{m_1,m_2,m_3,m_4\}}| \{m_1,m_2,m_3,m_4\} \in \pairs \} \\
\texttt{KnownActiveN} &= \land \{ a_{i,m} = \texttt{NodeFun}_m (n_{i,1},\dots,n_{i,\mu}) | \nodef_m \text{ is defined.} \} \\
\texttt{KnownActiveE} &= \land \{ b_{i,j,q,m} = \texttt{EdgeFun}_m(e_{i,j,q,1}, .., e_{i,j,q,\mu})
| \edgef_m \text{ is defined.} \} \\
\texttt{PresentCons} & = \texttt{PresentE} \land \texttt{PresentN} \land 
\texttt{PresentP} \land \texttt{KnownActiveE} \land \\
&  \texttt{KnownActiveN}
\end{align*}
We also collect the variables that are not set to true in $\texttt{PresentCons}$.
\begin{align*}
\texttt{AbsentELabel} &=
\{e_{i,j,q,m}|(i,M_1,j),...,(i,M_{q'},j) \in \edges \land 
0 < q \leq q' \land m \not\in M_{q} \}\\
\texttt{AbsentE} &= \{e_{i,j,q}|(i,M_1,j),...,(i,M_{q'},j) \in \edges \land 
q' < q \leq \pi \}\\
\texttt{AbsentNLabel} &= \{n_{i,m}| m \not\in \nlabel(i) \land i \in \nodes \}\\
\texttt{AbsentP} &= \{p_{\{m_1,m_2,m_3,m_4\}}| {\{m_1,m_2,m_3,m_4\}} \not\in \pairs \} \\
\texttt{UnknownActive} &=  \Land \{ a_{i,m} = \nodef_m (n_{i,1},\dots,n_{i,\mu}) | \nodef_m \text{ is undefined.} \} \union \\
&\quad\quad\quad  \{ b_{i,j,q,m} = \edgef_m(e_{i,j,q,1}, .., e_{i,j,q,\mu})
| \edgef_m \text{ is undefined.} \}
\end{align*}
We have defined $\texttt{AbsentELabel}$, $\texttt{AbsentE}$, $\texttt{AbsentN}$, and
$\texttt{AbsentP}$
as sets.
%
They will be converted into formulas depending
on the different usage in the synthesis problems.

% Since not all the activity functions are defined some of the functions
% appearing in $\texttt{ActiveE}$ and $\texttt{ActiveN}$ are
% uninterpreted function symbols.

>>>>

%\subsection{Boolean satisfiability of the search problem}
We translate the search problem into a Boolean satisfiability
problem and use SMT solvers to find the satisfying VTSs.
%
We will first present the variables used to encode the
VTSs and the properties.
%
Afterwards, we will present the constraints corresponding to the
properties.

@@

%if all the pairs of the present molecule are present on the same edge than all of them inhibit each other resulting in making all the molecules inactive, we term this process ``pairing inhibition".
%
The constraints for the variants can be given to a SAT or a SMT solver to find
VTSs that belong to the variants.
\subsection{Encoding for $\fqbf$ search problem}\label{enc:univ-cond}
%
\input{satqbf-table}
Recall that the specification for universal condition~\ref{sec:property} requires quantifier alternation. 
%
Most SAT solvers can not efficiently handle the quantifier in the formulas. 
% 
Although, it is possible to encode the universal condition as a SAT or SMT problem.
%
There is always a danger of a blow-up in the formula size, making it harder for the SAT solver to handle. 
%Which may blow up the formula size for the solver to solve.
% 
As an experiment, we have implemented a direct SMT version (encoding not presented) of the universal problem extending the current $\fsmt$ search encodings.
%
The basic idea is to conjunct the formula for each combination value of the quantified variable. 
%
The table~\ref{tab:satqbf-graph}, shows the search results of this encoding for the universal condition with 4 nodes and 3-connected graphs. 
%
The connectivity constraint and connectedness constraints are tagged with subscript $U$, $\texttt{StabilityC}_{U}$, $\texttt{Connected}(k)_{U}$,  to distinguish from earlier presented encodings of existential condition.
%
The SMT extended version was not able to scale it to the desired results and either timed-out (T/O) or ran out of memory (M/O).
%
This can be due to the fact that directly translating a quantified formula into a quantifier-free  by introducing a vector of variables for each boolean possibility will blow up the size of the formula and may not be solvable by a solver.
%
%
\subsubsection{$\fqbf$ search encoding of the universal condition:}
We will describe several constraints that encode universal condition based on QBF formulation in this section.
%
To avoid cumbersome notation, we will not explicitly write the ranges
of the indexing in the constraints.
%
$i$ and $j$ will range over nodes, i.e., from $1$ to $\nu$.
%
$m$ will range over molecules, i.e., from $1$ to $\mu$.
%
$q$ will range over edges between two nodes, i.e., from $1$ to $\pi$.
%

Using constraints for VTS from Section~\ref{enc:smt}, we encode the basic consistency of VTSs. 
%
%Recall that the molecule set is divided into 3:1 Q and R pairs (or partition).
\begin{align*}
%\texttt{EdgeC} =\;&\hspace{-1ex}\bigwedge\limits_{i,j,q} (\bigvee_m e_{i,j,q,m} )\limplies e_{i,j,q}
%\land
%\bigwedge\limits_{i,q} \neg e_{i,i,q}
%\land
%% \tag{\texttt{NodeC}}\label{eq:f0}\\
%\bigwedge\limits_{\mathclap{i,j,q,m}} e_{i,j,q,m} \limplies (n_{i,m} \land n_{j,m} )
%\\
%\texttt{ActivityC} =\;&
%\bigwedge\limits_{\mathclap{i,j,q,m}} b_{i,j,q,m} \limplies e_{i,j,q,m} \quad\land\quad
%\bigwedge\limits_{i,m} a_{i,m} \limplies n_{i,m}
%% \tag{V2}\label{eq:f1}\\
%\\
%\texttt{PairingC} =\;&
%\exists qr. \Land\limits_{\mathrlap{m_1,m_2,m_3,m_4}\;\;}
%(p_{\{m_1,m_2,m_3,m_4\}} \limplies qr_{m_1} + qr_{m_2} + qr_{m_3} + qr_{m_4} = 3) 
%% \land\\ 
%%   & \quad\quad\quad ( m_1 = m_2 = m_3 = m_4  \limplies \lnot p_{m_1,m_2,m_3,m_4} )
%% \bigwedge\limits_{m} 
%% (\Lor\limits_{m'} p_{mm'} \limplies \Land_{m'} \lnot p_{m'm} ) \land
%% (\Lor\limits_{m'} p_{m'm} \limplies \Land_{m'} \lnot p_{mm'} )
%\\
%\texttt{Fusion1} =\;&
%\bigwedge\limits_{i,j,q} e_{i,j,q} \limplies
%\bigvee_{{m_1,m_2,m_3,m_4}} (\Land_{l=1}^4 ( b_{i,j,q,m_l} \lor a_{j,m_l} ) \land 
%\Lor\limits_{l=1}^4 b_{i,j,q,m_l} \land\\
%& \qquad \qquad \qquad \qquad \Lor\limits_{l=1}^4 a_{j,m_l}\land p_{\{m_1,m_2,m_3,m_4\}})
%\\
%\texttt{Fusion2} =\;&
%\bigwedge\limits_{\mathclap{i,j,q,m_1,m_2,m_3, l \in \{1,..,3\}}} b_{i,j,q,m_1} \land ..\land b_{i,j,q,m_l} \limplies \\
%& \hspace{2cm} \neg 
%\bigvee_{\mathclap{j \neq j^{\prime}, m_{l+1}^{\prime},..,m_{4}^{\prime}}} ( a_{j^{\prime},m_{l+1}^{\prime}} \land .. \land a_{j^{\prime},m_4^{\prime}} \land p_{\{m_1,..,m_l,m^{\prime}_{l+1},..,m^{\prime}_4\}})
\texttt{Consistancy} =\;& \texttt{EdgeC} \land
\texttt{ActivityC} \land \texttt{PairingC} \land
\texttt{Fusion} 
\end{align*}
%\texttt{EdgeC} states that each edge has at least one molecule,
%there are no self loops, and edge labels are consistent with node labels.   
%\texttt{ActivityC} states that active molecule are present.
%\texttt{PairingC} states that all molecules are divided into two types
%using $qr_m$ bit, which encodes if $m$ belongs to one type or another,
%and any fusing set of molecules must have three molecules involved
%from one type and one molecule from the other.
%\texttt{Fusion1}, and \texttt{Fusion2} states the well-fused condition.
%\texttt{Consistancy} is the conjunction of all of the above.

\paragraph{Activity functions}
%
Activity functions are encoded as presented in the Section~\ref{enc:smt}.
%We also need to encode that the activity of the molecules are
%controlled by activity functions.
%
%The input VTS may include concrete activity functions for some molecules,
%and for the others the functions may be unknown and to be synthesized. 
%
%The concrete functions can be given to us in many different ways,
%for example as a lookup table, or a concise Boolean formula.
%
%In the following section, we will assume the appropriate encoding is
%used for the concrete functions and represent them by \texttt{NodeFun}$_m$
%and \texttt{EdgeFun}$_m$ for node and edge regulations respectively.
%
%We will use $\nodef_m$ and $\edgef_m$ to represent functions that
%are unknown in a VTS.
%
% \begin{align}
% \bigwedge\limits_{i,m} n_{i,m} \limplies a_{i,m} =  \nodef_m (n_{i,1},\dots,n_{i,\mu}) 
% \tag{\texttt{ActiveN}}\label{eq:anb}\\
%    \bigwedge\limits_{i,j,q,m} e_{i,j,q,m} \limplies b_{i,j,q,m} = \edgef_m(e_{i,j,q,1}, .., e_{i,j,q,\mu} )
%   \tag{\texttt{ActiveE}}\label{eq:aeb}
% \end{align}
%
%Later we will be synthesizing the unknown activity functions and 
%replace $\nodef_m$ and $\edgef_m$ with parameterized constraints that
%encode a space of candidate functions.
%\subsubsection{VTS properties}
%For the synthesis of incomplete systems,
%we need properties against which we synthesize the missing parts.
%
Next, we will discuss stability and $k$-connectedness.
%

\paragraph{Stability property}
%
%We use Boolean variable $r_{i,j,m,p}$ to indicate if there is an
%$m$-path from $i$ to $j$ of length less than or equal to $p$.
%
%We use $m$-reachability to encode the stability condition in VTSs.
%%
%The following constraint recursively encodes that node $j$ is
%$m$-reachable from node $i$ in less than $p$ steps.
%%
We encode stability condition using the reachability variables like presented in Section~\ref{enc:smt}. 
%
Recall the definition of $\texttt{Paths}(r)$ and $\texttt{Loop}(r)$ constraint from Section~\ref{enc:smt}.
%  
\begin{align*}
\texttt{Paths}(r) &= \bigwedge\limits_{\mathclap{i,j,m,p}} r_{i,j,m,p} \limplies (\bigvee_{q} \, e_{i,j,q,m} \lor \bigvee_{i\neq i^{\prime}} ( \, \bigvee_{q} e_{i,i^{\prime},q,m}) \land r_{i^{\prime},j,m,p-1} )
\\
\texttt{Loop}(r) &= \bigwedge\limits_{i,j,m} (\bigvee_{q} e_{i,j,q,m}) \limplies r_{j,i,m,\nu}
\end{align*}
We encode stability, \texttt{QStability} using quantifiers as existence of reachability bits such that constraints $\texttt{Paths}(r)$ and $\texttt{Loop}(r)$ are satisfied.
\begin{align*}
\texttt{QStability} &= \exists r. \; \texttt{Paths}(r) \land \texttt{Loop}(r)
\end{align*}
\paragraph{$k$-connected property}
%
The $k$-connectedness expresses robustness against failure of few edges.
%
%Let us use $d_{i,j,q}$ to indicate $q$th edge between $i$ and $j$ is failed
%and $r'_{i,j}$ to indicate if there is a path from $i$ to $j$ in
%the modified VTS.
%
% To check whether $k$-connected is a
% necessary condition, we remove (drop) $k-1$ edges from the graph and
% if it disconnects the graph, and we get an assignment.
% %
% We have a graph that is not $k$-connected.
%
Recall the definition of $\texttt{Fail}(d,k)$, $\texttt{FReach}(d,r')$ and $\texttt{Connected}(r')$ constraint from Section~\ref{enc:smt}.
%In the following,  encodes that only
%existing edges can be failed and exactly $k-1$ edges are failed.
%%
% defines reachability in the modified VTS.
%%
%We use a new variable $r'_{i,j,p}$ to encode reachability from
%$i$ to $j$ in at most $p$ steps.
%%
%$\texttt{Connected}(r')$ says that all nodes are reachable from any
%other node.
\begin{align*}
\texttt{Fail}(d,k) = & 
\bigwedge\limits_{i,j,q} d_{i,j,q} \limplies e_{i,j,q}  \land 
\sum_{i,j,q} d_{i,j,q} = k-1\\
% \texttt{ReachAbove}(d,r') = &
% \bigwedge\limits_{i,j}  [\bigvee_{q} (e_{i,j,q} \land  \neg d_{i,j,q}) \lor  (\bigvee_{i' \neq i}  r^{\prime}_{i',j} \land  \bigvee_{q} (e_{i,i',q} \land \neg d_{i,i',q}) ] \limplies r^{\prime}_{i,j}\\
\texttt{FReach}(d,r') = &\hspace{-1ex}
\bigwedge\limits_{i,j,p}  \hspace{-1ex}r^{\prime}_{i,j,p} \hspace{-1ex}\limplies\hspace{-1ex} [\bigvee_{q} (e_{i,j,q} \land  \neg d_{i,j,q}) \hspace{-1pt}\lor \hspace{-2pt} (\bigvee_{\mathclap{i' \neq i}}  r^{\prime}_{i',j,p-1} \land  \bigvee_{q} (e_{i,i',q} \land \neg d_{i,i',q}) ]\\
\texttt{Connected}(r') = & \Land\limits_{i,j} (r^{\prime}_{i,j,\nu} \lor r^{\prime}_{j,i,\nu})
\end{align*}
%We will be synthesizing $k$-connected graphs.
%
We define $\texttt{QConnected}(k)$ that says for all possible valid failures
the graph remains reachable. 
\begin{align*}
\texttt{QConnected}(k) = & \forall d.\;
(\texttt{Fail}(d,k) \limplies \exists r'.\;\texttt{FReach}(d,r')
\land \texttt{Connected}(r'))
% \\
% \texttt{Disconnected}(k) = & \exists d.\;
%         (\texttt{Drop}(d,k) \land \exists r'.\;\texttt{ReachAbove}(d,r')
%                               \land \lnot \texttt{Connected}(r'))
\end{align*}
Since $d$ variables in $\texttt{QConnected}(k)$ are universally
quantified, $\texttt{QConnected}(k)$ introduces quantifier alternations.
%
%Therefore, synthesis against this property will require QBF reasoning.
%%
%We may make the formula quantifier free  by considering all possible failures
%separately and introducing a vector of reachability variables for each
%failure.
%%
%However, this will blow up the size of the formula and may not be
%solvable by a SAT solver.
%
%\begin{align*}
%  & \forall G \, \texttt{Connectivity(k)} \limplies  \exists
%                        f,p: \texttt{Stability} \, \land \, \texttt{Consistancy}  
%  \tag{U}\label{eq:univcond}
%\end{align*}

@@

In our synthesis method, we take a VTS $G =
(\nodes,\mols,\edges,\nlabel,\pairs,\edgef,\nodef)$ as input.
%
We allow activity functions not to be specified.
%
We will re-use the $\fqbf$ encodings for the input VTS as described in the Section~\ref{enc:univ-cond}.
%
We construct the following constraints to encode the available information
about $G$.
%
We encode both the present and the absent components in $G$.
%
Later, the constraints will help us encode the synthesis problems.
%
\begin{align*}
\texttt{PresentE} &= \land \{e_{i,j,q,m}|(i,M_1,j),...,(i,M_{q'},j) \in \edges \land q \leq q' \land m \in M_q \}\\
\texttt{PresentN} &= \land \{n_{i,m}| m \in \nlabel(i) \land i \in \nodes \}\\
\texttt{PresentP} &= \land \{p_{\{m_1,m_2,m_3,m_4\}}| \{m_1,m_2,m_3,m_4\} \in \pairs \} \\
\texttt{KnownActiveN} &= \land \{ a_{i,m} = \texttt{NodeFun}_m (n_{i,1},\dots,n_{i,\mu}) | \nodef_m \text{ is defined.} \} \\
\texttt{KnownActiveE} &= \land \{ b_{i,j,q,m} = \texttt{EdgeFun}_m(e_{i,j,q,1}, .., e_{i,j,q,\mu})
| \edgef_m \text{ is defined.} \} \\
\texttt{PresentCons} & = \texttt{PresentE} \land \texttt{PresentN} \land 
\texttt{PresentP} \land \texttt{KnownActiveE} \land \\
&  \texttt{KnownActiveN}
\end{align*}
We also collect the variables that are not set to true in $\texttt{PresentCons}$.
\begin{align*}
\texttt{AbsentELabel} &=
\{e_{i,j,q,m}|(i,M_1,j),...,(i,M_{q'},j) \in \edges \land 
0 < q \leq q' \land m \not\in M_{q} \}\\
\texttt{AbsentE} &= \{e_{i,j,q}|(i,M_1,j),...,(i,M_{q'},j) \in \edges \land 
q' < q \leq \pi \}\\
\texttt{AbsentNLabel} &= \{n_{i,m}| m \not\in \nlabel(i) \land i \in \nodes \}\\
\texttt{AbsentP} &= \{p_{\{m_1,m_2,m_3,m_4\}}| {\{m_1,m_2,m_3,m_4\}} \not\in \pairs \} \\
\texttt{UnknownActive} &=  \Land \{ a_{i,m} = \nodef_m (n_{i,1},\dots,n_{i,\mu}) | \nodef_m \text{ is undefined.} \} \union \\
&\quad\quad\quad  \{ b_{i,j,q,m} = \edgef_m(e_{i,j,q,1}, .., e_{i,j,q,\mu})
| \edgef_m \text{ is undefined.} \}
\end{align*}
We have defined $\texttt{AbsentELabel}$, $\texttt{AbsentE}$, $\texttt{AbsentN}$, and
$\texttt{AbsentP}$
as sets.
%
They will be converted into formulas depending
on the different usage in the synthesis problems. 
% Since not all the activity functions are defined some of the functions
% appearing in $\texttt{ActiveE}$ and $\texttt{ActiveN}$ are
% uninterpreted function symbols.
%
\subsubsection{Encoding synthesis property}
%
We will do synthesis against the following property that says the VTS
is stable and $k$-connected.
%
Where $k$ is the minimum connectivity constraint for the corresponding variant in the result of existential condition.
% The property is designed to balance the search space such that the synthesis procedure does not
% succeed with simply adding too many edges. 
%
\begin{align*}
\texttt{Property} =  \texttt{QStability} \land \texttt{QConnected}(k) 
%\land \texttt{DisConnected}(4)
\end{align*}
The property was proposed in~\cite{shukla2017discovering}.
%
However, the biological relevance of the property is debatable and open for change.
%
Our tool is easily modifiable to support any other property that may be deemed 
interesting by the biologists.
